Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-15 write-ups/PLDI/paper.tex
Related work stuff.

@@

There has been a lot of work on accelerating blocking irregular
algorithms using GPUs~\cite{...}, and on the \emph{persistent threads}
programming style for long-running kernels~\cite{...}.  These
approaches rely on the occupancy-bound execution model, and aim to
take ownership of the GPU by scheduling as many workgroups as the GPU
has compute units, relying on fair scheduling between these
workgroups.  This means the GPU is unavailable for other tasks, and
the method is not future-proof because (as discussed throughout the
paper) the occupancy-bound execution model is not guaranteed by GPU
programming models and is unlikely to be guaranteed in practice by
future GPU platforms.

@@

Hardware support for preemptive multitasking for \nvidia GPUs has been
proposed, enabling full preemption as well as \emph{SM-draining},
whereby workgroups occupying a symmetric multiprocessor (SM; a compute
unit using our terminology) are drained until the SM becomes free for
other tasks.  A follow-up work adds the notion of SM \emph{flushing},
whereby a workgroup running on an SM can be cancelled and re-scheduled
from scratch if the workgroup has not yet committed any visible
side-effects to shared memory.  Both approaches have been evaluated
using simulators, over sets of regular GPU kernels.

@@

\paragraph{Dynamic parallelism} Recent version of CUDA and OpenCL allow a kernel to spawn further kernels~\cite{...}.
This \emph{dynamic parallelism} can be exploited to implement a GPU-based scheduler: an initial scheduler kernel uses the feature to spawn additional kernels as required~\cite{ }.

 is launched

%
\ADComment{Kepler preemption support?  Mention in related work?}

\ADComment{Sreepathi mentioned the NCCL framework.}

>>>>

\begin{figure}
\includegraphics[width=\columnwidth]{cooperative_overhead.pdf}
\caption{Overhead: ratio of execution time}\label{fig:overhead}
\end{figure}

@@

There has been a lot of work on accelerating blocking irregular
algorithms using GPUs~\cite{...,TPO10}, and on the \emph{persistent threads}
programming style for long-running kernels~\cite{...}.  These
approaches rely on the occupancy-bound execution model, and aim to
take ownership of the GPU by scheduling as many workgroups as the GPU
has compute units, relying on fair scheduling between these
workgroups.  This means the GPU is unavailable for other tasks, and
the method is not future-proof because (as discussed throughout the
paper) the occupancy-bound execution model is not guaranteed by GPU
programming models and is unlikely to be guaranteed in practice by
future GPU platforms.

@@

Hardware support for preemptive multitasking for \nvidia GPUs has been
proposed, enabling full preemption as well as \emph{SM-draining},
whereby workgroups occupying a symmetric multiprocessor (SM; a compute
unit using our terminology) are drained until the SM becomes free for
other tasks.  A follow-up work adds the notion of SM \emph{flushing},
whereby a workgroup running on an SM can be cancelled and re-scheduled
from scratch if the workgroup has not yet committed any visible
side-effects to shared memory.  Both approaches have been evaluated
using simulators, over sets of regular GPU kernels.  State-of-the-art \nvidia GPUs targeting data centers (where multi-tasking is essential) do support
preemption, though it is not clear whether they guarantee fair scheduling~\cite{PascalWhitepaper}.

@@

\paragraph{Dynamic parallelism} Recent version of CUDA and OpenCL allow a kernel to spawn further kernels~\cite{CUDASpec,opencl2Spec}.
This \emph{dynamic parallelism} can be exploited to implement a
GPU-based scheduler, by having an initial scheduler kernel repeatedly
spawn further kernels as required, according to some scheduling
policy~\cite{DBLP:conf/ppopp/Muyan-OzcelikO16}.  However, a kernel
that uses dynamic parallelism is still subject to the scheduling
limitations of current GPU programming models, and in particular the
problem that workgroups are not fairly scheduled.  Hence this feature
does not offer a solution to the problem of deploying blocking
algorithms on GPUs in a safe manner that we tackle with this work.
